---
title: "Suicides In India-Kaggle Dataset"
author: "Albab Shams"
date: "08 April 2018"
output: html_document
---

```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(dplyr)
library(tidyr)
library(reshape2)
library(data.table)
```

```{r,warning=FALSE}
sui <- read.csv("D:/DataSets/Practice datasets/suicides_in_India.csv")

sui_cleaned <- sui[which(sui$State!="Total (All India)" & sui$State!="Total(Uts)" & sui$State!="Total (States)") & sui$Age_group!= "0-100+", ]
```

Grouping the data by "state" and "year"
```{r}
sui_1 <- sui_cleaned %>% group_by(State,Year) %>% summarise(total=sum(Total))
```

Creating a new varibale called "growth rate".This'll give the change in number of suicide cases YOY for each state.
```{r,warning=FALSE}
sui_2<- mutate(sui_1,gowth_rate_yoy=(total/lag(total)-1))

#Removing the details for Lakshadweep,as it contain 'inf' and 'NaN'
sui_2n <- sui_2[-c(217:228), ]
```
Generating line graph for YOY growth rate
```{r}
attach(sui_2n)
ggplot(aes(x=Year,y=gowth_rate_yoy*100),data=filter(sui_2n,State %in% c("Bihar","Kerala","West Bengal","Madhya Pradesh")))+geom_point(aes(color=State))+geom_line(aes(color=State))+scale_x_continuous(breaks = seq(2001,2012,1))+scale_y_continuous(breaks = seq(-50,100,20))+geom_hline(yintercept = 0,linetype=5)+ylab("YOY Growth(%)")
```

Grouping the data by "age group" and "gender"
```{r}
sui_3 <- sui_cleaned %>% group_by(Gender,Age_group) %>% summarise(Total_sui=sum(Total))
```
Generating the bar graph showing suicide counts in various age-groups

```{r}
ggplot(aes(x=sui_3$Age_group,y=sui_3$Total_sui),data = sui_3)+geom_bar(stat="identity",aes(fill=sui_3$Gender),position="dodge")+xlab("AGE_GROUPS")+ylab("No of Suicides")+ggtitle("Suicide by age_group")
```
Grouping the data by state and age group

```{r}
sui_4 <- sui_cleaned %>% group_by(State,Age_group) %>% summarise(total=sum(Total))
```
```{r}
ggplot(aes(x=sui_4$State,y=sui_4$total,fill=sui_4$Age_group),data = sui_4)+geom_bar(stat="identity",position = "dodge")+xlab("States")+ylab("Total suicides")+theme(axis.text.x = element_text(angle = -90,vjust=0,hjust=0))
```


Creating a data frame with columns as "year" and "total suicides"
```{r}
sui_YOY <- data.frame(tapply(sui$Total,sui$Year,sum))
sui_YOY <- cbind(rownames(sui_YOY),data.frame(sui_YOY),row.names=NULL)
names(sui_YOY) <- c("year","Total_Suicides")
sui_YOY <- transform(sui_YOY,year=as.numeric(levels(year))[year])
ggplot(data = sui_YOY,aes(x=year,y=Total_Suicides))+geom_line()+scale_x_continuous(breaks = seq(2001,2012,1))
```

Analyzing the causes of suicides
```{r}
sui_causes <- sui[which(sui$Type_code=="Causes"), ]
causes_summary <- sui_causes %>% group_by(Type,Gender) %>% summarise(SUM=sum(Total))
ggplot(data=causes_summary,aes(x=causes_summary$Type,y=causes_summary$SUM))+geom_bar(stat = "identity",aes(fill=causes_summary$Gender),position ="dodge")+theme(axis.text.x=element_text(angle =-90,vjust = 0,hjust=0))+ggtitle("Suicide Causes-by Gender")+theme(plot.title = element_text(hjust=.5,family="serif"))+xlab("Suicide Causes")+ylab("Suicide count")
```
